Truthfulness Verification System

### 2011-02-23 trunk/proposal/paper.tex
fixed sections and subsections

\section{Introduction}

@@

%We are currently working to get the T-verifier code running on our
%systems, with inputs from Xian Li. So the descriptions below lacks some required
%details and makes some assumptions.
%
%
%As of now T-Verifier produces 5 alternative statements once it is provided with a
%doubtful statement (DS), a doubtful unit (DU), data type of the unit etc in the
%first phase. Though these sentences produced are quite accurate, the facility to
%pinpoint the correct alternative does exist yet. The project goal, may be formulated as:
%
%"Extract information from Wikipedia and ontologies ( such as Yago, DBpedia and
%Freebase) which helps to verify truthfulness of statements. Describe the
%extraction algorithm and how the information can be utilized to assist statement
%truthfulness verification."

%figure

@@

\subsection{Freebase} 
Freebase has information about approximately 20 million Topics,
each one having a unique Id, which can help distinguish multiple entities which
have similar names, such as Henry Ford the industrialist vs Henry Ford the
footballer. Most of the topics are associated with one or more types (such as
people, places, books, films, etc) and may have additional properties like "date
of birth" for a person or latitude and longitude for a location.

Python library: http://code.google.com/p/freebase-python/ The type system:
http://blog.freebase.com/2007/03/19/the-type-system-in-freebase/

Freebase seems to be the first point to start not only because of
the richness of the data available but also the available api.

\subsection{Dbpedia}

http://wiki.dbpedia.org/Datasets?v=lbf
http://wiki.freebase.com/wiki/DBPedia
http://sourceforge.net/projects/sparql-wrapper/files/sparql-wrapper-python/1.4.2/

A few major tasks would be:

\subsection{Yago}
YAGO is a semantic knowledge base with over 900,000 entities (like persons,
organizations, cities, etc.). It knows about 6 million facts about these entities

\section{Truthfulness verification }
\subsection {Building queries from the data supplied by the user}

@@

From what we have explored till now, by quering Freebase we can get the
various property, value pairs of the entity that we have supplied in the form
of $TU$. In case the result is $\phi$, it suggests that we do not have
information for this particulat combination $TU$ and $AU_{i}$. Due to the
richness of the information in Freebase we can assume with a considerable degree of confidence
that this combination is untrue. This is one way verification, but in case of
non-null results, how do we filter and rank the results with the available set
of information is something we can not formulate at this stage with out further
exploration.

\subsection{Application of Machine learning algorithms}: 
Though application of machine learning algorithms like using Naive Bayesian,
neural network

\section{2.1 Supervised - Use a naive bayesian classifier}
\section{Unsupervised methods - clustering or Probabilistic LSI}
\section{Partially Supervised methods}

>>>>

\section{Introduction}
Web has became most prevalent source of information now a days. However, many
information on the web is untruthful. Also because of the widespread reach of
web, sometimes web is used to propagate untruthful facts for social and political
reasons. Hence with the increasingly use of web as information source,
verification of information truthfulness became an important facet to consider.

The popular search engines extract information from web based on keywords and
metadata without considering the truthfulness of the facts. Also, there have been
not much research in this area. To our best knowledge, the most recent work on this area is T-verifier  \cite{tverifier}, which uses results from search engines
to verify truthfulness of statements. T-verifier performs very well on the
test-dataset. However, T-verifier has some problems with the overall approach to
verify the truthfulness of statements and there is room for improvement. In this
work, we will extend the T-verifier system so that it's weaknesses can be
resolved to get a more robust truthfulness verification system.

@@

% We are currently working to get the T-verifier code running on our systems,
% with inputs from Xian Li. So the descriptions below lacks some required details
% and makes some assumptions.   As of now T-Verifier produces 5 alternative
% statements once it is provided with a doubtful statement (DS), a doubtful unit
% (DU), data type of the unit etc in the first phase. Though these sentences
% produced are quite accurate, the facility to pinpoint the correct alternative
% does exist yet. The project goal, may be formulated as:  "Extract information
% from Wikipedia and ontologies ( such as Yago, DBpedia and Freebase) which helps
% to verify truthfulness of statements. Describe the extraction algorithm and how
% the information can be utilized to assist statement truthfulness verification."

\section{Problems with Current System}
Although results from the tested dataset achieves good performance (90\%
accuracy), failing of T-verifier for some statements shows that it has some
inherent problems and there is room for improvement.

First, T-verifier assumes that truthful statements will be more propagated in the
web compared to the untruthful statement. However, this may be not true in
general because of intended and planned propaganda for establishing some
untruthful statements. This kind of propagandas are even becoming more common now
a days due to widespread reach of Internet. T-verifier also showed that "Hillary
Clinton is the President of United States" has more hits than "Hillary Clinton is
the Secretary of State". Although T-verifier was able to find the correct
statement in this case using multiple ranks together, in general the untruthful
statement can be prevalent in the web compared to truthful statements.

Second, T-verifier do not use the reputation of the information source.
T-verifier uses only search results returned by the search engine irrespective of
the origin and believability of the information origin. Hence there is room for
improvement here to give more weight to the information obtained from trustworthy
sources such as Wikipedia.

@@

\subsubsection{Freebase} 
Freebase has information about approximately 20 million Topics, each one having a
unique Id, which can help distinguish multiple entities which have similar names,
such as Henry Ford the industrialist vs Henry Ford the footballer. Most of the
topics are associated with one or more types\cite{freebasetype} (such as
people, places, books, films, etc) and may have additional properties like "date of birth" for a person
or latitude and longitude for a location. Freebase not only contains data from
the Wikipedia but also other sources; users can submit data to the Freebase
datastore and expand it in richness. We tinkered with the api\cite{freebaseapi}
and it appeared to be the most viable starting point for the project.

\subparagraph{Motivational Example}

\begin{lstlisting}[label=some-code,caption=Minimal code to Freebase]

import freebase
import pprint

query = [{
  "a:starring": [{
    "actor": "Meg Ryan"                                                       
  }],
  "b:starring": [{                                                            
    "actor": "Tom Hanks"
  }],
  "type": "/film/film",
  "*": [],
}]

pp = pprint.PrettyPrinter(indent=4)
result = freebase.mqlread(query)

print "Movie names & their various forms"

for i in result:                                                              
        pp.pprint(i["key"])

\end{lstlisting}
\begin{lstlisting}[label=output,caption=Cleaned Output]

Movie names & their various forms
[   '158982',
    'You$0027ve_Got_Mail',
    '18171032',

     ...
    'E-m$0040il_f$00FCr_Dich',
    'youve-got-mail']
[   '176489',
    'Joe_Versus_the_Volcano',
    'Joe_Vs$002E_The_Volcano',
    'Brain_Cloud',
     ...
    '2327353',
    'joe-versus-the-volcano']
[   '226198',
    'Sleepless_in_Seattle',
    'Sleepless_In_Seattle',
     ....
    '169146',
    '106482',
    'Insonnia_d$0027amore',
    '62812',
    'Schlaflos_in_Seattle',
    'sleepless-in-seattle'\]

\end{lstlisting}

The above results show how the three movies starring Tom Hanks and Meg Ryan.
When we query Google with Tom Hanks and Meg Ryan, the top result is a page from
Answers.com where a user has asked which are the movies where the two actors
appear together, and the answer lists these three movies namely - ``Joe versus
the Volcano'', `` Sleepless in Seatle'' and ``You\'ve got Mail''. A quick lookup
of the Wikipedia and IMDB pages also confirm the same.

\subsubsection{Dbpedia}
DBpedia is a similar project to Freebase, but it focuses mainly on the content
available from Wikipedia. It scores in being precisely importing the data from
the info boxes in Wikipedia pages, but at this stage it does not seem to be
offering anything additional over Freebase \cite{freebasedbpedia}. We are yet to
explore its programmatic interface \cite{dbpediaapi}.

\subsubsection{Yago}
YAGO is a semantic knowledge base with over 900,000 entities (like persons,
organizations, cities, etc.) and uses Wikipedia and Wordnet as its main source of
information. We are yet to explore the programmatic interfaces it provides and
how we can use it for the project.

\subsection{Truthfulness verification }
\subsubsection {Building queries from the data supplied by the user}

@@

From what we have explored till now, by quering Freebase we can get the
various property, value pairs associated with an entity which is the $TU$ in
our case. In case the result is $\phi$, it suggests that we do not have
information for this particulat combination $TU$ and $AU_{i}$. Due to the richness of the information in Freebase we can assume with a considerable degree of confidence
that this combination is untrue. This is one way verification, but in case of
non-null results, how do we filter and rank the results with the available set
of information is something we can not formulate at this stage with out further
exploration.

\subsubsection{Application of Machine learning algorithms}
Though usage of machine learning algorithm seems to be very exciting approach at
this point, we are also skeptical about it. Supervised classifiers might not be
benficial for us due to the limited dataset for training and the potentialy
unlimited variety of inputs to classify. However, if we can use Graph
Based Partially Supervised Learning\cite{conf/acl/CelikyilmazTH09} or
Spreading Activation \cite{Cohen1987255} on the information extacted from the
rich ontologies is something we plan to explore.
